Learning to Synthesize Rules for Network Intrusion Detectors

### 2020-01-08 paper/paper.tex
pass on examples
comments
pass on 3 and 4.1

\subsection{Denial-of-Service}

@@

Finally, \tname{} ranks the inferred rules according to heuristics
extracted from a database of existing rules. One example heuristic
function we used to rank the generated rules was
\Luc{
%\begin{wrapfigure}[14]{r}{0.5\textwidth}
\begin{figure}[H]
  \centering
%  \vspace{-7ex}
  \scalebox{0.85}{
    \begin{tikzpicture}
      \begin{axis}[
          ybar stacked,
          enlargelimits=0.15,
          legend style={at={(0.5,-0.15)},
            anchor=north,legend columns=-1},
          %          ylabel={\#rules},
          symbolic x coords={0, 1, 2, 3, 4, 5, 6, 7, 8, 9, 10},
          xtick=data,
          xticklabel style={rotate=45},          
          nodes near coords,
          every node near coord/.append style={font=\scriptsize},        
          nodes near coords align={vertical},
        ]
        \addplot coordinates {(0,2134) (1,9164) (2,5103) (3,5841) (4,1987) (5,1622) (6,805) (7, 535) (8, 340) (9,154) (10, 53)};
%        \legend{relevant, irrelevant}
      \end{axis}
    \end{tikzpicture}
  }
  \vspace{-2ex}
  \caption{\label{fig:distribution-rules-protocol}Number of rules per number of contents.}
%\end{wrapfigure}
\end{figure}

}

 \Fix{...informa um
  exemplo concreto que vc. usaram...}. The list below shows the rule
options for the top-5 rules \tname{} produces for this attack.

@@

\subsection{Active Reconnaissance}

@@

\begin{figure}[h!]
  \lstinputlisting[language=C,numbers=none,keywords={dsize,itype}]{pingscan.suricata}
  \caption{Suricata rule for Ping Scan.}
  \label{fig:pingscan-example}
\end{figure}

@@

\tname{} is a technique to synthesize rules for signature-based
\nids~(\eg{}, Snort~\cite{snort} and Suricata~\cite{suricata}). The
goal of \tname\ is to create rules that capture \emph{only} the
malicious traffic.

\begin{figure*}[t]
\centering
\includegraphics[trim=0 350 100 0,clip,width=0.65\textwidth]{figs/nids-workflow}
\caption{\tname\ workflow.}
\label{fig:overview}
\end{figure*}

\vspace{1ex}
\noindent\textbf{Overview.}~To create rules, \tname\ leverages (1)
benign and malicious network traffic and (2) rules from public
rulesets. Figure~\ref{fig:overview} shows the workflow of \tname{} as
a pipeline of three components, which appear numbered in the figure.
The inputs of the technique appear at the bottom of the figure whereas
the output appears at the right. \tname\ takes on input malicious and
benign traffic and a set of rules from public rulesets and produces on
output a list of plausible rules to isolate the malicious traffic.

@@

The goal of the first step is to extract options from the positive
traffic. For that, \tname{} parses the protocol message, looking for
fields in the message associated with options (see
Section~\ref{sec:rules-and-packets}).  Unfortunately, inferring
options at the field granularity level is insufficient to capture
\CodeIn{content} options (see
Section~\ref{sec:example-suricata-rules}), which refer to parts of the
payload of the message and cannot be extracted directly from the
fields. Handling this case is critically important as \Fix{xx}\% of the
rules contain \CodeIn{content} options as the one from
Section~\ref{sec:content-example}.

\pgfplotsset{width=6cm,compat=1.8}
\pgfplotsset{every tick label/.append style={font=\tiny}}

%\begin{wrapfigure}[15]{r}{0.56\textwidth}
\begin{figure}[t!]
  \centering
%  \vspace{-4ex}
  \scalebox{1.2}{  
    \begin{tikzpicture}
      \begin{axis}[
          bar width=6pt,
          scaled ticks=false,
          tick label style={/pgf/number format/fixed},
          ybar stacked,
          enlargelimits=0.15,
          legend style={at={(0.5,-0.15)},
            anchor=north,legend columns=-1},
          symbolic x coords={0, 1, 2, 3, 4, 5, 6, 7, 8, 9, 10},
          xtick=data,
          nodes near coords,
          every node near coord/.append style={font=\tiny},
          nodes near coords align={vertical},
        ]
        \addplot coordinates {(0,25) (1,10222) (2,4956) (3,5589)
          (4,1651) (5,1457) (6, 729) (7, 471) (8, 313) (9, 137) (10, 126)};
      \end{axis}
    \end{tikzpicture}
  }
  \caption{\label{fig:distribution-contents}Histogram of number of
    \CodeIn{content} options per rule (up to 10).}
%\end{wrapfigure}
\end{figure}

Figure~\ref{fig:distribution-contents} shows the histogram of the
number of these options per rule (up to 10) obtained from a public
ruleset~\cite{emerging-threats-open} containing 27.8K Suricata
rules. Rules without \CodeIn{content} options is an exception and half
of the rules contain at least two \CodeIn{content} options.  To find
the values associated with these options, \tname{} splits the payload
of the malicious message in tokens using natural language delimiters,
such as $\backslash$t, $\backslash$n, $\backslash$r, and spaces. We
observed, by inspecting existing rules, that \CodeIn{content} options
typically refer to sequence of characters in the payload of messages
separated by these delimiters. This tokenization step can generate
dozens of options. For the example from
Section~\ref{sec:content-example}, \tname{} produces a total of
\Fix{xx} content options.

\Mar{pessoal, isto explica como inferimos opcoes de content, mas nao
  explica como inferimos as outras opcoes. Em particular, as opcoes
  dos primeiros dois exemplos da Secao 3.}

\Mar{-------------------------------- estou aqui}

@@

@@

>>>>

\subsection{Denial-of-Service}
\label{sec:dos}

@@

Finally, \tname{} ranks the inferred rules according to heuristics
extracted from a database of existing rules. One example heuristic
function we used to rank the generated rules was
\Fix{...elaborate...}. The list below shows the rule options for the
top-5 rules \tname{} produces for this attack.

@@

\subsection{Active Reconnaissance}
\label{sec:active-recon}

@@

\begin{figure}[h!]
  \vspace{1ex}
  \lstinputlisting[language=C,numbers=none,keywords={dsize,itype}]{pingscan.suricata}
  \caption{Suricata rule for Ping Scan.}
  \label{fig:pingscan-example}
  \vspace{1ex}  
\end{figure}

@@

\tname{} is a technique to synthesize rules for signature-based
\nids~(\eg{}, Snort~\cite{snort} and Suricata~\cite{suricata}). The
goal is to create rules that capture \emph{only} the
malicious traffic.

\begin{figure*}[t!]
\centering
\includegraphics[trim=0 350 100 0,clip,width=0.65\textwidth]{figs/nids-workflow}
\caption{\tname\ workflow.}
\label{fig:overview}
\end{figure*}

\vspace{1ex}
\noindent\textbf{Overview.}~To create rules, \tname\ leverages (1)
benign and malicious network traffic and (2) rules from public
rulesets. \tname\ takes on input malicious and benign traffic and a
set of rules from public rulesets and produces on output a list of
plausible rules to isolate the malicious
traffic. Figure~\ref{fig:overview} shows the workflow of \tname{} as a
pipeline of three components, which appear numbered in the figure.
The inputs of the technique appear at the bottom of the figure whereas
the output appears at the right.

@@

The goal of the first step is to extract options from the positive
traffic, which is characterized with pcap files~\cite{pcap}, a popular
format to describe network packets.

For that, \tname{} parses the message looking for options associated
with the data. As described in Section~\ref{sec:rules-and-packets},
there is a correspondence between packet fields and rule
options. \tname{} leverages that mapping to infer options. The
examples of sections~\ref{sec:dos} and~\ref{sec:active-recon} show
cases where \tname{} infers options with this mapping. Unfortunately,
inferring options at the field granularity level is insufficient to
capture \CodeIn{content} options as the described on
Section~\ref{sec:content-example}. These options refer to parts of the
payload of the message and cannot be extracted directly from the
fields. Handling this case is critically important as \Fix{xx}\% of
the rules contain \CodeIn{content} options.

\pgfplotsset{width=6cm,compat=1.8}
\pgfplotsset{every tick label/.append style={font=\tiny}}
\begin{figure}[t!]
  \centering
  \scalebox{1.2}{  
    \begin{tikzpicture}
      \begin{axis}[
          bar width=6pt,
          scaled ticks=false,
          tick label style={/pgf/number format/fixed},
          ybar stacked,
          enlargelimits=0.15,
          legend style={at={(0.5,-0.15)},
            anchor=north,legend columns=-1},
          symbolic x coords={1, 2, 3, 4, 5, 6, 7, 8, 9, 10},
          xtick=data,
          nodes near coords,
          every node near coord/.append style={font=\tiny},
          nodes near coords align={vertical},
        ]
        \addplot coordinates {(1,10222) (2,4956) (3,5589)
          (4,1651) (5,1457) (6, 729) (7, 471) (8, 313) (9, 137) (10, 126)};
      \end{axis}
    \end{tikzpicture}
  }
  \caption{\label{fig:distribution-contents}Histogram of number of
    \CodeIn{content} options per rule.}
\end{figure}

Figure~\ref{fig:distribution-contents} shows the histogram of the
number of these options per rule obtained from a public
ruleset~\cite{emerging-threats-open} containing 27.8K Suricata
rules. We limited the number of content options per rule to 10 for
space. Rules without \CodeIn{content} options is an exception
(\Fix{xx}\%) and half of the rules contain at least two
\CodeIn{content} options.  To find the values associated with these
options, \tname{} splits the payload of the malicious message in
tokens using natural language delimiters, such as $\backslash$t,
$\backslash$n, $\backslash$r, and spaces. We observed, by inspecting
existing rules, that \CodeIn{content} options typically refer to
sequence of characters in the payload of messages separated by these
delimiters. This tokenization step can generate dozens of options.
\tname{} produced a total of \Fix{xx} content options for the example
Section~\ref{sec:content-example}.

@@

\Mar{-------------------------- estou aqui}

@@

\subsection{Step 3: Ranking}

\Fix{...elaborate...}

\begin{figure}[H]
  \centering
  \scalebox{0.85}{
    \begin{tikzpicture}
      \begin{axis}[
          ybar stacked,
          enlargelimits=0.15,
          legend style={at={(0.5,-0.15)},
            anchor=north,legend columns=-1},
          %          ylabel={\#rules},
          symbolic x coords={0, 1, 2, 3, 4, 5, 6, 7, 8, 9, 10},
          xtick=data,
          xticklabel style={rotate=45},          
          nodes near coords,
          every node near coord/.append style={font=\scriptsize},        
          nodes near coords align={vertical},
        ]
        \addplot coordinates {(0,2134) (1,9164) (2,5103) (3,5841) (4,1987) (5,1622) (6,805) (7, 535) (8, 340) (9,154) (10, 53)};
      \end{axis}
    \end{tikzpicture}
  }
  \vspace{-2ex}
  \caption{\label{fig:distribution-rules-protocol}Number of rules per number of contents.}
\end{figure}
